---
title: "threadleR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{threadleR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, include = FALSE}
library(threadleR)
can_run_threadle <- th_is_available()
```

## Overview

threadleR is an R client for [Threadle](https://threadle.dev), a high-performance network engine. It starts a background Threadle process, sends commands, and returns results back to R. Most functions are thin wrappers around Threadle CLI commands.

## Installation
Install from GitHub:
```{r eval=FALSE}
# install.packages("remotes")
remotes::install_github("YukunJiao/threadleR")
```

Load the package:
```{r eval=can_run_threadle}
library(threadleR)
```

## Prerequisites

### Check whether Threadle is available

`th_is_available()` checks whether the `threadle` executable can be found on your system.

```{r eval=FALSE}
th_is_available()
```

If it returns FALSE, either add threadle to your PATH or start Threadle with an explicit path:

```{r eval=FALSE}
th_start_threadle("/full/path/to/threadle")
```

### Start Threadle
Start the Threadle background process:
```{r eval=can_run_threadle}
th_start_threadle()
```

### Stop Threadle
Stop the Threadle process when you are done:
```{r eval=can_run_threadle}
th_stop_threadle()
```

## Working directory
Threadle has its own working directory. Many file-based commands (especially `th_load_file()`) depend on it.

To inspect Threadle working directory:
```{r eval=FALSE}
th_get_workdir()
```

To set it:
```{r eval=FALSE}
th_set_workdir("~documents")   # or "~", or an absolute path
```

To keep Threadle aligned with the current R working directory:
```{r eval=FALSE}
th_sync_wd()
```

## Example data: Lazega

Threadle network files may reference companion files (e.g., a nodeset file).
When you load a network from disk, Threadle will also try to load the referenced nodeset file using its current working directory.

### Stage example files to your working directory
```{r eval=can_run_threadle}
th_start_threadle()

th_sync_wd()
exdir <- th_stage_examples_to_wd(folder = "threadle_examples", overwrite = TRUE)

th_set_workdir(exdir)

list.files(exdir)
```

### Load Lazega network (and its referenced nodeset)
```{r eval=can_run_threadle}
net_file <- file.path(exdir, "lazega.tsv")

lazega <- th_load_file("lazega", net_file, type = "network")

# th_load_file("lazega", ..., type = "network") also creates:
# lazega_nodeset  (a threadle_nodeset handle in the calling environment)
th_i()
th_info(lazega)
th_info(lazega_nodeset)
```

### Preview and inventory
```{r eval=can_run_threadle}
th_preview(lazega)
th_i()
```

## Adding and removing ties

### One-mode layer: edges
Create a layer and add a few edges:
```{r eval=can_run_threadle}
th_add_layer(lazega, "friends2", mode = 1, directed = FALSE, valuetype = "binary")

th_add_edge(lazega, "friends2", node1id = 1, node2id = 2)
th_add_edge(lazega, "friends2", node1id = 2, node2id = 3)

th_check_edge(lazega, "friends2", node1id = 1, node2id = 2)
th_get_all_edges(lazega, "friends2", offset = 0, limit = 10)
```

Remove an edge or clear a layer:
```{r eval=can_run_threadle}
th_remove_edge(lazega, "friends2", node1id = 1, node2id = 2)
th_clear_layer(lazega, "friends2")
```

### Two-mode layer: affiliations (hyperedges)
```{r eval=can_run_threadle}
th_add_layer(lazega, "clubs", mode = 2)

th_add_hyper(lazega, "clubs", hypername = "group1", nodes = c(1, 2, 3))
th_get_node_hyperedges(lazega, "clubs", nodeid = 1)
th_get_hyperedge_nodes(lazega, "clubs", hypername = "group1")

th_add_aff(lazega, "clubs", nodeid = 4, hypername = "group1")
th_remove_aff(lazega, "clubs", nodeid = 4, hypername = "group1")

th_get_all_hyperedges(lazega, "clubs", offset = 0, limit = 10)
```

## Nodes and attributes
### Add nodes
Add a node to the nodeset:
```{r eval=can_run_threadle}
th_add_node(lazega_nodeset, nodeid = 999)
th_get_nbr_nodes(lazega_nodeset)
```

### Define and query attributes
Define an integer attribute and set values:
```{r eval=can_run_threadle}
th_define_attr(lazega_nodeset, "score", "int")
th_set_attr(lazega_nodeset, nodeid = 1, attrname = "score", attrvalue = 10)
th_get_attr(lazega_nodeset, nodeid = 1, attrname = "score")
th_get_attr_summary(lazega_nodeset, "score")
```
Remove values or undefine the attribute:
```{r eval=can_run_threadle}
th_remove_attr(lazega_nodeset, nodeid = 1, attrname = "score")
th_undefine_attr(lazega_nodeset, "score")
```

Generate random attributes:
```{r eval=can_run_threadle}
th_generate_attr(lazega_nodeset, "x", attrtype = "int", min = 1, max = 10)
th_get_attr_summary(lazega_nodeset, "x")
```

## Basic network routines
Compute density (for a one-mode layer):
```{r eval=can_run_threadle}
th_density(lazega, "friends")
```
Degree (writes an attribute):
```{r eval=can_run_threadle}
th_degree(lazega, "friends", attrname = "deg", direction = "both")
th_get_attr(lazega, nodeid = 2, attrname = "deg")
```

Connected components (writes component IDs to an attribute):
```{r eval=can_run_threadle}
th_components(lazega, "friends", attrname = "comp")
th_get_attr_summary(lazega, "comp")
```

Shortest path:
```{r eval=can_run_threadle}
th_shortest_path(lazega, node1id = 1, node2id = 3, layername = "friends")
```


## Import and export layer data
Export a layer to an edge list:
```{r eval=can_run_threadle}
out <- tempfile(fileext = ".tsv")
th_export_layer(lazega, "friends", file = out, header = TRUE, sep = "\t")
readLines(out, n = 3)
unlink(out)
```
Import edges into an existing layer:
```{r eval=can_run_threadle}
tmp <- tempfile(fileext = ".tsv")
write.table(
  data.frame(node1 = c(1, 2), node2 = c(2, 3), value = c(1, 1)),
  file = tmp, sep = "\t", row.names = FALSE, col.names = FALSE, quote = FALSE
)

th_import_layer(lazega, "friends",
  file = tmp, format = "edgelist",
  node1col = 0, node2col = 1, valuecol = 2,
  header = FALSE, sep = "\t"
)

th_check_edge(lazega, "friends", node1id = 1, node2id = 2)
unlink(tmp)
```

## Save and load Threadle structures
Save nodeset or network:

```{r eval=can_run_threadle}
tmp_net <- tempfile(fileext = ".tsv")
th_save_file(lazega, file = tmp_net)
file.exists(tmp_net)
unlink(tmp_net)
```
Load back from file:
```{r eval=can_run_threadle}
# Ensure Threadle working directory can resolve referenced files if loading a network
# (for a standalone roundtrip, prefer staging all required files into one folder first)
```

## Cleanup

Delete individual structures:
```{r eval=can_run_threadle}
th_delete(lazega)
th_delete(lazega_nodeset)
th_i()
```

Or delete everything:
```{r eval=can_run_threadle}
th_delete_all()
th_i()
```

Stop Threadle:

```{r eval=can_run_threadle}
th_stop_threadle()

if (exists("exdir", inherits = FALSE) && is.character(exdir) && nzchar(exdir) && dir.exists(exdir)) {
  unlink(exdir, recursive = TRUE, force = TRUE)
}
```

## Session info
```{r session-info}
sessionInfo()
```
