---
title: "threadleR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{threadleR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, include = FALSE}
library(threadleR)
can_run_threadle <- th_is_available()
```

## Overview

threadleR is an R client for [Threadle](https://threadle.dev), a high-performance network engine. It starts a background Threadle process, sends commands, and returns results back to R. The public R functions use a `th_` prefix and wrap the commands shown by Threadle's `help` command.

## Installation

Install from GitHub:

```{r eval=FALSE}
# install.packages("remotes")
remotes::install_github("YukunJiao/threadleR")
```

Load the package:

```{r eval=FALSE}
library(threadleR)
```

## Start Threadle

Check whether the `threadle` executable is available:

```{r eval=FALSE}
th_is_available()
```

Start a Threadle process. If `threadle` is not on your PATH, pass the full path to the executable.

```{r eval=can_run_threadle}
th_start_threadle()
```

```{r eval=FALSE}
th_start_threadle("/full/path/to/threadle")
```

## Load example data

The easiest way to use the bundled data is `th_load_examples()`. It loads the example files shipped with threadleR, sets Threadle's working directory to the example-data folder, and assigns R handles such as `mynet` and `mynet_nodeset`.

```{r eval=can_run_threadle}
examples <- th_load_examples("mynet")
names(examples)

th_i()
th_info(mynet)
th_info(mynet_nodeset)
```

You can also inspect the Threadle working directory:

```{r eval=can_run_threadle}
th_get_workdir()
th_dir()
```

## Explore nodes and attributes

```{r eval=can_run_threadle}
th_get_nbr_nodes(mynet)
th_get_all_nodes(mynet, offset = 0, limit = 8)
th_get_nodeid_by_index(mynet, index = 0)
th_get_random_node(mynet)
```

The `mynet` example has a `gender` node attribute:

```{r eval=can_run_threadle}
th_get_attr(mynet_nodeset, nodeid = 123, attrname = "gender")
th_get_attrs(mynet_nodeset, nodes = c(123, 234, 345), attrname = "gender")
th_get_attr_summary(mynet_nodeset, attrname = "gender")
```

## One-mode layers

Use one-mode layers such as `kinship` and `trade` for edge and degree commands:

```{r eval=can_run_threadle}
th_get_edge(mynet, "trade", node1id = 123, node2id = 345)
th_check_edge(mynet, "kinship", node1id = 123, node2id = 345)
th_get_all_edges(mynet, "trade", offset = 0, limit = 10)
th_get_random_edge(mynet, "trade")
```

```{r eval=can_run_threadle}
th_get_node_alters(mynet, nodeid = 345, layernames = "trade", direction = "both")
th_get_degree(mynet, nodeid = 345, layernames = "trade", direction = "both")
th_get_random_alter(mynet, nodeid = 345, layernames = "trade", direction = "both")
th_density(mynet, "kinship")
```

Some routines write results back as node attributes:

```{r eval=can_run_threadle}
th_components(mynet, "kinship", attrname = "kinship_component")
th_degree(mynet, "kinship", attrname = "kinship_degree", direction = "both")
th_get_attr_summary(mynet, "kinship_degree")
```

Shortest-path commands use the public `layernames` argument:

```{r eval=can_run_threadle}
th_shortest_path(mynet, node1id = 123, node2id = 567, layernames = "kinship")
paths_by_gender <- th_shortest_paths(
  "paths_by_gender",
  mynet,
  attrname = "gender",
  layernames = "kinship"
)
th_info(paths_by_gender)
```

## Two-mode layers

The `work` layer in `mynet` is a two-mode layer:

```{r eval=can_run_threadle}
th_get_all_hyperedges(mynet, "work", offset = 0, limit = 10)
th_get_hyperedge_nodes(mynet, "work", hypername = "ias")
th_get_node_hyperedges(mynet, "work", nodeid = 123)
```

You can add and remove affiliations with the exported two-mode helpers:

```{r eval=can_run_threadle}
th_add_hyper(mynet, "work", hypername = "liu_temp", nodes = c(123, 234))
th_add_aff(mynet, "work", nodeid = 345, hypername = "liu_temp")
th_get_hyperedge_nodes(mynet, "work", hypername = "liu_temp")

th_remove_aff(mynet, "work", nodeid = 345, hypername = "liu_temp")
th_remove_hyper(mynet, "work", hypername = "liu_temp")
```

## Derived structures

Create a filtered nodeset and subnet from the example data:

```{r eval=can_run_threadle}
female_nodeset <- th_filter(
  "female_nodeset",
  mynet_nodeset,
  attrname = "gender",
  cond = "eq",
  attrvalue = "f"
)
female_mynet <- th_subnet("female_mynet", mynet, female_nodeset)

th_info(female_nodeset)
th_info(female_mynet)
```

Layer transformations also operate on the example network:

```{r eval=can_run_threadle}
th_symmetrize(mynet, "trade", method = "max", newlayername = "trade_sym")
th_dichotomize(
  mynet,
  "trade",
  cond = "ge",
  threshold = 1000,
  truevalue = 1,
  falsevalue = 0,
  newlayername = "trade_hi"
)
th_project_two_mode(mynet, "work", method = "count", newlayername = "work_projected")

th_get_edge(mynet, "trade_sym", 123, 345)
th_get_edge(mynet, "trade_hi", 890, 234)
th_get_edge(mynet, "work_projected", 123, 456)
```

## Editing structures

You can create new structures when you need a small scratch network:

```{r eval=can_run_threadle}
scratch_nodes <- th_create_nodeset("scratch_nodes", name = "Scratch nodes", createnodes = 5)
scratch_net <- th_create_network("scratch_net", scratch_nodes, name = "Scratch network")

th_add_layer(scratch_net, "friendship", mode = 1, directed = FALSE, valuetype = "binary")
th_add_edge(scratch_net, "friendship", node1id = 1, node2id = 2)
th_check_edge(scratch_net, "friendship", node1id = 1, node2id = 2)

th_remove_edge(scratch_net, "friendship", node1id = 1, node2id = 2)
th_clear_layer(scratch_net, "friendship")
```

Node and attribute commands use the same public handles:

```{r eval=can_run_threadle}
th_add_node(scratch_nodes, nodeid = 10)
th_get_all_nodes(scratch_nodes, offset = 0, limit = 20)
th_remove_node(scratch_nodes, nodeid = 10)

th_define_attr(scratch_nodes, "score", "int")
th_set_attr(scratch_nodes, nodeid = 1, attrname = "score", attrvalue = 7)
th_get_attr(scratch_nodes, nodeid = 1, attrname = "score")
th_remove_attr(scratch_nodes, nodeid = 1, attrname = "score")
th_undefine_attr(scratch_nodes, "score")
```

## Import and export

```{r eval=can_run_threadle}
out <- tempfile(fileext = ".tsv")
th_export_layer(mynet, "kinship", file = out, header = TRUE, sep = "\t")
readLines(out, n = 3)
unlink(out)
```

```{r eval=can_run_threadle}
tmp <- tempfile(fileext = ".tsv")
write.table(
  data.frame(node1 = c(1, 2), node2 = c(2, 3), value = c(4, 5)),
  file = tmp,
  sep = "\t",
  row.names = FALSE,
  col.names = FALSE,
  quote = FALSE
)

th_add_layer(scratch_net, "imported_edges", mode = 1, directed = FALSE, valuetype = "valued")
th_import_layer(
  scratch_net,
  "imported_edges",
  file = tmp,
  format = "edgelist",
  node1col = 0,
  node2col = 1,
  valuecol = 2,
  header = FALSE,
  sep = "\t"
)
th_get_edge(scratch_net, "imported_edges", node1id = 1, node2id = 2)
unlink(tmp)
```

Save and load Threadle's internal file formats:

```{r eval=can_run_threadle}
tmp_nodes <- tempfile(fileext = ".tsv")
th_save_file(scratch_nodes, file = tmp_nodes)
scratch_nodes_copy <- th_load_file("scratch_nodes_copy", file = tmp_nodes, type = "nodeset")
th_get_nbr_nodes(scratch_nodes_copy)
unlink(tmp_nodes)
```

## Low-level commands

`th_cmd()` is an escape hatch for Threadle commands that do not yet have a typed wrapper.

```{r eval=can_run_threadle}
th_cmd("info", args = list(structure = mynet))
```

## Cleanup

```{r eval=can_run_threadle}
th_delete_all()
th_i()
th_stop_threadle()
```

## Session info

```{r session-info}
sessionInfo()
```
